Profiles of Wind Direction and Speed in Turbulent Ekman Flow

### 2021-07-29 write/manuscript.tex
proceeded on manuscript

\usepackage[margin=3.5cm]{geometry}
\usepackage{lineno}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{xcolor}

@@

\usepackage[maxcitenames=2,style=authoryear-comp,firstinits=true,citestyle=authoryear,natbib=true,backend=bibtex]{biblatex}
\DeclareNameAlias{sortname}{family-given}
\renewcommand*{\multinamedelim}{\addcomma\space}
\renewcommand*{\finalnamedelim}{~and\space}
\setlength{\bibitemsep}{\baselineskip}
\renewbibmacro{in:}{}

@@

\newcommand{\todo}[1]{\textcolor{red}{$[$#1$]$}}

@@

Wind veer away from its geostrophic direction is
(i) due to direct frictional effects in the very vicinity of the surface and
(ii) due to turbulence which exerts indirect frictional effects; these effects
cause a slow-down of the mean wind reducing the Coriolis force thus turning the wind
in favor of the pressure gradient force. 
%%
Not only does the veering set the frame of reference for surface layer theory, it also has
physical effects on large and small scales.
%
From a large-scale perspective, the veering of wind across the planetary boundary layer determines
the amount of cross-isobaric mass-flux, commonly referred to as 'Ekman pumping' \citep{ekman:AMA1905},
and it is thus a key factor in the life-cycle of large-scale synoptic systems.
%
% %% This shear is probably not related to Ekman-turning 
% On the mesoscale, shear in the surface layer a crucial factor for the initiation and sustainment of strong convective events (REFERENCE?).
% 
Locally, the directional shear of the wind in the upper part of the surface layer may cause a systmatic yaw
for tall wind power generation devices where blades reach into the Ekman layer, i.e. that part of the boundary
layer where the wind starts to turn. 
%
\par
%
Wind veer in the planetary boundary layer is commonly characterized by the surface veering angle $\alpha$
defined as the angle between the negative surface shear stress $\tau_\SFC$ and the geostrophic wind.
%
Surface veering $\alpha$ and geostrophic drag $Z\equiv u_\star/G$, where $u_\star\equiv\sqrt{\nu|\tau_\SFC|}$,
uniquely determine the surface drag $\tau_\SFC$ in a turbulent Ekman flow. 
%
In any quantitative description of the surface layer, the friction velocity $u_\star$ is the dynamic scale
and $\alpha$ defines the alignment of the frame of reference.
%
Knowledge about $u_\star$ and $\alpha$ is thus a pre-requisite for any quantitative theory of the surface layer and a first
attempt in theoretical quantification was undertaken by \citet{rossby:PPO1935} using integral relations in the boundary layer.
%
Asymptotic similarity theory was later used by \citet{tennekes:JAS1973, blackadar:JAS1968},
and--based on his seminal direct numerical simulations (DNS) of Ekman flow--,~\cite{spalart:JFM1989} suggested
a modification to take into account effects of low to intermediate Reynolds numbers.
Later on, constants were re-evaluated with a focus on the atmospheric boundary layer based on
observations \cite{hogstrom:BM1988,hogstrom:BM1996}
and numerical modelling \citep{spalart:PF2008, spalart:PF2009, ansorge:BM2014, ansorge:BM2019}. 
%
\par
%
Attempts were also undertaken to obtain profiles of the wind speed: 
One approach is to match the inner and outer layer at a reference height;
\citet{etling:2002, emeis:2018}  (Sec. 21.10; Eq. 21.48) choose the Prandtl-layer height to match
the wind speed profiles, which, however, requires to externally prescribe the veering $\alpha$ at that height. 
A one-dimensional profile with constant veering is given by \citet[Sec. 3; Eq. 3.1-3.19]{emeis:m2007}.

@@

\citet{jiang:JAS2018} recognized that the outer part of the Ekman boundary layer receives less attention in comparison

@@

Universal profiles of the wind vector for turbulent Ekman flow can serve as well-described limit and reference
for higher-order approaches taking into account possible effects of stratification, roughness or other physical
complications encountered in the realgeophysical system.

@@

Such generalized profiles may also serve as better initial condition for numerical simulation of the flow,

@@

relies on sub-grid closures that often assume alignment of the stress with gradients -- a questionable pre-requisite when the wind rotates with height.

@@

Despite advances in analysis of this simplified set-up \citep{jiang:JAS2018} \todo{(MORE?)},

@@

arbitrary external forcing (manifest in variation of the Reynolds number).

@@

for such simulation.

@@

Here, the routinely employed concept of Reynolds-number similarity can help: it states that certain statistics (normally those of low order) become independent of the Reynolds number when scaled appropriately \todo{Dimotakis / da Silva / Pope referece?} .

@@

It appears that for certain statistics in Ekman flow, this regime is reached with the Reynolds numbers that became
possible with the increasing capabilities of supercomputering over the past years.

@@

The robust features of mean velocity profiles across a range of Reynolds numbers
are exploited in this work to develop a formulation for
both the stream-wise and span-wise components of the mean velocity vector as a function of the
Reynolds number.

\todo{Numerical approach, DNS, Scale-separation, Re-similarity
  \citep{dimotakis:JFM2000, dasilva:ARF2014, pope:2000, moin:ARF1998, mellado:JAM2018}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Problem Formulation and Numerical Approach}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
We consider here incompressible, turbulent Ekman flow, that is, the turbulent flow over a flat rotating plate, as
a physical model for the truly neutral atmospheric boundary layer (ABL).
%
The f-plane approximation is applied such that rotation only acts on horizontal velocity components;
we thus neglect rotational effects on the horizontal components of velocity and dynamical effects
due to latitudinal variation of the rate of rotation.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Notation and Governing Equations}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
The dimensional velocity vector of the numerical simulations is $\underline{U} = (U_1,U_2,U_3) = (U,V,W)$ over the coordinate system
$Oxyz$, where an approximate alignment (plus/ minus few degrees centigrade) of the direction $Ox$ with the surface shear stress is achieved.
The coordinate $Oy$ points away from the wall,  and $Oz$ points in the span-wise direction normal to $Oxy$.
For analysis of the results, we use two coordinate systems that are
(i)  exactly aligned with the surface shear stress and labelled by an upper index $\alpha$ as in $\underline{U}^\alpha$ for the velocity vector, and
(ii) the coordinate system aligned with the free-atmosphere geostrophic wind labelled by an upper index $G$ as in $\underline{U}^G$.
%
We denote by $u_\star$ the modulus surface friction velocity and let $Z_\star=1/u_\star$ for brevity. 
%
\par
%
We consider the incompressible Navier--Stokes equations for the three velocity components on the f-plane in a framework
that is governed by the physical parameters
(i)   geostrophic wind magnitude $G=\sqrt{G_1^2+G_3^2}$ (where $\mathbf{G}=(G_1,0,G_3)$ is the geostrophic wind vector),
(ii)  Coriolis parameter $f$ (representing the angular rotation), and
(iii) kinematic viscosity $\nu$.
   %
These yield a Reynolds number $\RE$ as non-dimensional parameter governing the system.
%
We use the geostrophic wind as velocity and the Coriolis parameter $f$ as time scale for the non-dimensional framework which implies the Rossby radius $\LR=G/f$ as length scale, such that
\begin{equation}
  \RE=\frac{G \LR }{\nu}. 
\end{equation}
%
\par
%
The non-dimensionalized set of governing equation reads as 
\begin{subequations} 
\begin{eqnarray}
  \frac{\partial u_i}{\partial t} &=& \frac{\p \pi}{\p x_i} - u_j \frac{\p u_i}{\p x_j} + \epsilon_{i2j} (u_j -G_j)   + \frac{1}{\RE} \frac{\p^2 u_i}{\p x_{j}^2} \\ 
  \frac{\partial u_j}{\partial x_j} &=& 0,   
\end{eqnarray} 
\label{eqn:governing} 
\end{subequations}
where $u_i=U_i/G$ are the non-dimensional components velocity, $\pi$ is non-dimensional pressure, and $g_j=G_j/G$ are non-dimensionalized components geostrophic wind (with $g_{1}^2+g_{3}^2=1$ by construction). 
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Numerical Simulations}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\todo{DNS Code}
\par
\todo{(new and old) Simulations}

@@

\end{itemize}
\begin{figure}
  \centerline{\includegraphics[width=0.7\textwidth]{../plot/ustar_alpha.pdf}}
  \caption{Variation of $Z_\star$ and surface veering with Reynolds number according to
    the theory by Spalart et al. (1989) and as estimated from DNS data}
  \label{fig:drag_law}
\end{figure}
% 
\subsection{Drag-Law}
A drag-law for Ekman flow determines two key parameters of the flow as a function of Reynolds number alone.
%
These are, first, the modulus of the surface friction normalized by the geostrophic wind forcing, $u_\star$ and,
second, the direction of surface shear stress, $\alpha$, also termed wind veering. 
%
A non-zero veering of the wind is a rather special case in comparison with most turbulent flows conisdered in an
engineering context, and it confronts us with a situation where the most appropriate coordinate system for analysis
(namely that aligned with the surface shear stress) is a priori unknown. 
%
We compare our DNS data against a semi-empirical drag-law based on integral consideration \cite{spalart:JFM1989}
and find, as demonstrated in previous work \cite{ansorge:BM2014}, excellent agreement in the range
 $400<\mathbf{Re}<1600$, representing a factor of 16 in variation of viscosity. 
%
\par
%
We also find that the solution of the transient equation involved in estimation of $u_\star$ for a given Reynolds number
$Re_D$ is approximated reasonably by the formulation
\begin{equation}
  u_\star= 4\log(\mathbf{Re}_D))-8
\end{equation}
which quantifies the 'weak' dependence of $u_\star$ on the Reynolds number as an approximately logarithmic one, at least
for problems with a scale separation on the order that is relevant to geophysical problem ($Re_D<10^{8}$).  
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\subsection{Profile in the outer layer}
%
The outer layer of Ekman flow is characterized by a turning of the wind velocity and the super-geostrophic maximum
that is sustained by momentum convergence at the inflection point of the velocity profile. 
The super-geostrophic maximum of streamwise velocity and a secondary minimum aloft the bulk-turbulent part of
the boundary layer are well-described by a classic Ekman spiral with adapted boundary conditions and a shift
in reference height.
%
The textbook solution for the Ekman layer reads as 
\todo{equations / figures}
\begin{figure}
  \includegraphics[width=\textwidth]{../plot/outer_layer.pdf}
  \caption{} 
\end{figure}

@@

\paragraph{Inner Layer.} It is important to note that the logarithmic layer observed here for the spanwise velocity does not

>>>>

\usepackage[margin=3.5cm]{geometry}
\usepackage{lineno}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{xcolor}
\usepackage{dirtytalk}
\usepackage{colortbl}
\usepackage{booktabs}

@@

\usepackage[maxcitenames=2,style=authoryear-comp,firstinits=true,citestyle=authoryear,natbib=true,backend=bibtex]{biblatex}
\DeclareNameAlias{sortname}{family-given}
\renewcommand*{\multinamedelim}{\addcomma\space}
\renewcommand*{\finalnamedelim}{~and\space}
\setlength{\bibitemsep}{\baselineskip}
\setlength\parindent{3em}
\setlength\parskip{0.5em}
\renewbibmacro{in:}{}

@@

\newcommand{\todo}[1]{\textcolor{red}{$[$#1$]$}}
\newcommand{\quotation}[1]{"#1"}

@@

Wind veer away from its geostrophic direction is
(i) due to direct frictional effects in the very vicinity of the surface and
(ii) due to turbulence which exerts indirect frictional effects; these effects
cause a slow-down of the mean wind reducing the Coriolis force thus turning the wind
in favor of the pressure gradient force. 
%%
Not only does the veering set the frame of reference for surface layer theory, it also has
effects at small and large scales.
%
From a large-scale perspective, the veering of wind across the planetary boundary layer determines
the amount of cross-isobaric mass-flux, commonly referred to as 'Ekman pumping' \citep{ekman:AMA1905},
and it is thus a key factor in the life-cycle of large-scale synoptic systems.
%
% %% This shear is probably not related to Ekman-turning 
% On the mesoscale, shear in the surface layer a crucial factor for the initiation and sustainment of strong convective events (REFERENCE?).
% 
At the small-scale end, the directional shear of the wind in the upper part of the surface layer
may cause a systmatic yaw for tall wind power generation devices where blades reach into the Ekman layer,
i.e. that part of the boundary layer where the wind starts to turn. 
%
\par
%
Wind veer in the planetary boundary layer is commonly characterized by the surface veering angle $\alpha$
defined as the angle between the negative surface shear stress $\tau_\SFC$ and the geostrophic wind.
%
Surface veering $\alpha$ and geostrophic drag $Z\equiv u_\star/G$, where $u_\star\equiv\sqrt{\nu|\tau_\SFC|}$,
uniquely determine the surface drag $\tau_\SFC$ in a turbulent Ekman flow. 
%
In any quantitative description of the surface layer, the friction velocity $u_\star$ is the dynamic scale
and $\alpha$ defines the alignment of the frame of reference.
%
Knowledge about $u_\star$ and $\alpha$ is thus a pre-requisite for any quantitative theory of the surface layer and a first
attempt in theoretical quantification was undertaken by \citet{rossby:PPO1935} using integral relations in the boundary layer.
%
Asymptotic similarity theory was later used by \citet{tennekes:JAS1973, blackadar:JAS1968},
and--based on his seminal direct numerical simulations (DNS) of Ekman flow--,~\cite{spalart:JFM1989} suggested
a modification to take into account effects of low to intermediate Reynolds numbers.
Later on, constants were re-evaluated with a focus on the atmospheric boundary layer based on
observations \citep{hogstrom:BM1988,hogstrom:BM1996}
and numerical modelling \citep{spalart:PF2008, spalart:PF2009, ansorge:BM2014, ansorge:BM2019}. 
%
\par
%
Attempts were also undertaken to obtain profiles of the wind speed: 
One approach is to match the inner and outer layer at a reference height;
\citet{etling:2002, emeis:2018}  (Sec. 21.10; Eq. 21.48) choose the Prandtl-layer height to match
the wind speed profiles, which, however, requires to externally prescribe the veering $\alpha$ at that height. 
A one-dimensional profile with constant veering is given by \citet[Sec. 3; Eq. 3.1-3.19]{emeis:m2007}.

@@

Still, in \citeyear{jiang:JAS2018}, \citeauthor{jiang:JAS2018}
recognized that the outer part of the Ekman boundary layer receives less attention in comparison

@@

Universal profiles of the wind vector for turbulent Ekman flow not only are a well-described limit
for theoretical exploration or higher-order approaches taking into account possible effects of stratification,
roughness or other physical complications encountered in the realgeophysical system.

@@

Furthermure, profiles can serve as better initial condition for numerical simulation of the flow,

@@

relies on sub-grid closures that often assume alignment of the stress with gradients -- a questionable
pre-requisite when the wind rotates with height.

@@

\citet{esau:EFM2004} investigated the representation of the Ekman boundary layer by dynamical subgrid closures
and \cite{zikanov:JFM2003} proposed a closure for the wind profile using a linearized representation of the eddy viscosity.
Despite advances in analysis of this simplified set-up \citep{jiang:JAS2018}, 

@@

arbitrary external forcing (manifest in variation of the Reynolds number) -- and indeed the fundamental
questions pertaining to such relatively simple dynamics of turbulence are not reflected in the
research foci of LES for the ABL over the past 50 years \citep{stoll:BM2020}. 

@@

for such simulation \citep{dimotakis:ARF2005}.

@@

Here, the routinely employed concept of Reynolds-number similarity can help. It postulates the existence
of \say{\emph{fully developed turbulence}  believed to occur for a sufficiently large but finite Reynolds number}
\citep{barenblatt:PF1995}.

@@

(Already in \citeyear{moin:ARF1998}, this in fact lead \citeauthor{moin:ARF1998} to the question \emph{how high a Re is high enough}?) 

@@

In fully developed turbulence certain statistics such as dissipation \citep{dimotakis:ARF2005} or profiles of
mean velocity \citep{barenblatt:JFM1993} become independent of the Reynolds number when scaled appropriately while
other statistics, such as the near-wall maximum in velocity fluctuation depend on $\RE$ \citep{baars:JFM2020a} and external
interfaces may exert an impact on near-wall scaling \citep{dasilva:ARF2014}. 
It appears that for certain statistics in Ekman flow, fully-developed turbulence is reached with the
Reynolds numbers that became possible with the increasing capabilities of supercomputering over
the past decades.
This paper exploits the robust features of mean velocity profiles from direct numerical
simulation across a range of Reynolds numbers to formulate both the stream-wise and span-wise components
of the mean velocity vector as a function of the Reynolds number.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Problem Formulation and Numerical Approach}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider here incompressible, turbulent Ekman flow, that is, the turbulent flow over a flat rotating plate, as
a physical model for the truly neutral atmospheric boundary layer (ABL).
The f-plane approximation is applied such that rotation only acts on horizontal velocity components;
we thus neglect rotational effects on the horizontal components of velocity and dynamical effects
due to latitudinal variation of the rate of rotation.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Notation and Governing Equations}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The dimensional velocity vector of the numerical simulations is $\underline{U} = (U_1,U_2,U_3) = (U,V,W)$ over the coordinate system
$Oxyz$, where an approximate alignment (plus/ minus few degrees centigrade) of the direction $Ox$ with the surface shear stress is achieved.
The coordinate $Oy$ points away from the wall,  and $Oz$ points in the span-wise direction normal to $Oxy$.
For analysis of the results, we use two coordinate systems that are
(i)  exactly aligned with the surface shear stress and labelled by an upper index $\alpha$ as in $\underline{U}^\alpha$ for the velocity vector, and
(ii) the coordinate system aligned with the free-atmosphere geostrophic wind labelled by an upper index $G$ as in $\underline{U}^G$.
We denote by $u_\star$ the modulus surface friction velocity and let $Z_\star=1/u_\star$ for brevity. 
We consider the incompressible Navier--Stokes equations for the three velocity components on the f-plane in a framework
that is governed by the physical parameters
(i)   geostrophic wind magnitude $G=\sqrt{G_1^2+G_3^2}$ (where $\mathbf{G}=(G_1,0,G_3)$ is the geostrophic wind vector),
(ii)  Coriolis parameter $f$ (representing the angular rotation), and
(iii) kinematic viscosity $\nu$.
These yield a Reynolds number $\RE$ as non-dimensional parameter governing the system.
We use the geostrophic wind as velocity and the Coriolis parameter $f$ as time scale for the non-dimensional framework which implies the Rossby radius $\LR=G/f$ as length scale, such that
  \RE=\frac{G \LR }{\nu}. 
The non-dimensionalized governing equations read as 
  \frac{\partial u_i}{\partial t} &=& \frac{\p \pi}{\p x_i} - u_j \frac{\p u_i}{\p x_j} + \epsilon_{i2j} (u_j -G_j)   + \frac{1}{\RE} \frac{\p^2 u_i}{\p x_{j}^2} \\ 
  \frac{\partial u_j}{\partial x_j} &=& 0,   
\label{eqn:governing} 
where $u_i=U_i/G$ are the non-dimensional components velocity, $\pi$ is non-dimensional pressure, and $g_j=G_j/G$ are non-dimensionalized components
geostrophic wind (with $g_{1}^2+g_{3}^2=1$ by construction).
These equations are solved inside a bounded cube of size $L_x \times L_y \times L_z$ with periodic boundary conditions in the lateral directions $Oxz$, a no-slip--no-penetration boundary at $y=0$ and
a no-penetration, free-slip boundary at $y=L_y$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Numerical Simulations}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The problem is solved numerically by tLab\footnote{https://github.com/turbulencia/tlab},
an open-source tool-suite to simulate and analyze turbulent flows.
We use here a fourth-order--five-step Runge--Kutta integration and sixth-order compact schemes
for spatial derivatives in all directions.
The incompressibility constraint is enforced by a fractional step approach where the Poisson equation
for the pressure field is soled to machine accuracy using a combined spectral/compact approach as
described in \citet{mellado:ZaM2012}.
Simulations used here are shown in Tab.~\ref{tab:sim-setup}.
We extend an existing set of simulations for $\RE\in\{125\,000, 281\,250,\ 500\,000\}$ \citep[gray shading; cf.][]{ansorge:BM2014,ansorge:JFM2016} 
by new simulations at higher Reynolds numbers up to $1.28\times10^6$.
In total, this yields one order of magnitude variation in terms of the scale separation in the boundary layer.
\todo{(new and old) Simulations}
\begin{table}
  \caption{\label{tab:sim-setup} Direct numerical simulation datasets used in this work}
  \begin{tabular}{c r r c c c c c}
    \toprule 
    $\RE_\Lambda$ & $\RE_D$ & $L_{xy}/\Lambda$ & $N_x\times N_y\times N_z$ & $\Delta x^+$ & $\Delta y^+$ & $\left.\Delta z^+\right|_{z=0}$ \\ 
    \midrule
    \rowcolor{gray!30}
       $125\,000$ &    $500$& 1.08 & $2048\times2048\times192$&4.1& 4.1& 1.05\\
    \rowcolor{gray!30}
       $281\,250$ & $   750$& 1.08 & $3072\times3072\times384$&5.6&11.2& 1.60\\
    \rowcolor{gray!30}
       $500\,000$ & $1\,000$& 1.08 & $3072\times6144\times512$&9.3& 4.7& 1.14\\
    \midrule
       $845\,000$ & $1\,300$& 0.54 & $2560\times5120\times640$&8.9& 4.5& 0.99\\

    $1\,280\,000$ & $1\,600$& 0.54 & $3860\times7680\times960$&8.6& 4.3& 1.00\\ 
    \bottomrule 
  \end{tabular} 
\end{table}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overview of the new simulations at $\RE_\tau$ up to 3000}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
Uniqueness and necessity of new simulations
%
\par
%
PARAGRAPH ON BULK PARAMETERS AND SCALING (cf. \ref{tab:bulk_results})
\begin{table}
  \caption{Bulk characterization of the simulations \label{tab:bulk_results}}
  \begin{tabular}{r r r r r r r }
    \toprule
    Re & $\delta^+=\RE_\tau$& $u_\star$ & $\alpha$ & $\delta_{95}/\delta$ & $f u_\star^{-3}\int_{0}^{\delta} e \mathrm{d} z $ & $ G^{-3}\int_{0}^{\delta} \epsilon \mathrm{d}z$\\
      \midrule 
      \rowcolor{gray!30} 
      500 & 479 & 0.0619& 25.5 & 0.66& 0.88& 1.31 \\
      \rowcolor{gray!30} 
      750 & 886 & 0.0561& 21.0 & 0.65& 0.90& 1.34 \\
      \rowcolor{gray!30}
      1000& 1403& 0.0530& 18.8 & 0.62& 0.92& 1.30 \\
      \midrule
      1300& 2122& 0.0501& 17.9 & 0.59& 0.85& 1.24 \\
      1600& 2978& 0.0482& 17.2 & 0.61& 0.91& 1.21 \\ 
      \bottomrule
    \end{tabular}
\end{table} 
%
\par
%
PARAGRAPH ON VELOCITY AND STRESS PROFILES cf. \ref{fig:stresses}
\begin{figure}
  \includegraphics[width=0.66\textwidth]{../plot/eddy_viscosity.pdf}
  \caption{Reynolds stresses and eddy viscosity across the Ekman layer plotted in ineer scaling for
    various Reynolds number
    \label{fig:stresses}} 
\end{figure} 
%
\par
%
PARAGRAPH ON TURBULENCE STRUCTURE (2D SLICES, SPECTRA?) cf. Fig.~\ref{fig:slices}

\begin{figure}
  \centerline{\includegraphics[width=0.33\textwidth]{../plot/PLANES/tke_xz_10.pdf}
  \includegraphics[width=0.33\textwidth]{../plot/PLANES/tke_xz_50.pdf}
  \includegraphics[width=0.33\textwidth]{../plot/PLANES/tke_xz_400.pdf}}
  \centerline{\includegraphics[width=\textwidth]{../plot/PLANES/tke_xy_1.pdf}}
  \caption{horizontal slices of turbulence kinetic energy in the Buffer layer (i=10), Logarithmich Layer (i=100) and outer lzayer (i=400); coloring between percentiles 4 and 96 of the respective image. Lower panel:streamwise--vertical intersect through the domain} 
  \label{fig:slices} }

@@

\end{itemize}
\begin{figure}
  \centerline{\includegraphics[width=0.7\textwidth]{../plot/ustar_alpha.pdf}}
  \caption{Variation of geostrophic drag, $Z_\star$, and surface veering, $\alpha$, with Reynolds number according to
    the theory by Spalart et al. (1989) and as estimated from DNS data}
  \label{fig:drag_law}
\end{figure}
% 
\subsection{Drag-Law}
A drag-law for Ekman flow determines two key parameters of the flow as a function of Reynolds number alone.
%
These are, first, the modulus of the surface friction normalized by the geostrophic wind forcing, $u_\star$ and,
second, the direction of surface shear stress, $\alpha$, also termed wind veering. 
%
A non-zero veering of the wind is a rather special case in comparison with most turbulent flows conisdered in an
engineering context, and it confronts us with a situation where the most appropriate coordinate system for analysis
(namely that aligned with the surface shear stress) is a priori unknown. 
%
We compare our DNS data against a semi-empirical drag-law based on integral consideration \cite{spalart:JFM1989}
and find, as demonstrated in previous work \cite{ansorge:BM2014}, excellent agreement in the range
 $400<\mathbf{Re}<1600$, representing a factor of 16 in variation of viscosity. 
%
\par
%
We also find that the solution of the transient equation involved in estimation of $u_\star$ for a given Reynolds number
$Re_D$ is approximated reasonably by the formulation
\begin{equation}
  u_\star= 4\log(\mathbf{Re}_D))-8
\end{equation}
which quantifies the 'weak' dependence of $u_\star$ on the Reynolds number as an approximately logarithmic one, at least
for problems with a scale separation on the order that is relevant to geophysical problems ($Re_D<10^{8}$).  
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\subsection{Profile in the outer layer}
%
The outer layer of Ekman flow is characterized by a turning of the wind velocity and the super-geostrophic maximum
that is sustained by momentum convergence at the inflection point of the velocity profile. 
The super-geostrophic maximum of streamwise velocity and a secondary minimum aloft the bulk-turbulent part of
the boundary layer are well-described by a classic Ekman spiral with adapted boundary conditions and a shift
in reference height.
%
\begin{figure}
  \includegraphics[width=0.5\textwidth]{../plot/outer_layer_u.pdf}\includegraphics[width=0.5\textwidth]{../plot/outer_layer_w.pdf}
  \caption{} 
\end{figure}

@@

\subsub{Inner Layer.} It is important to note that the logarithmic layer observed here for the spanwise velocity does not
